### 2023-02-14 doc/manuscript/translocation.qmd
Moved 2 figures to supplement and fixed typos

Emerging infectious diseases pose a severe threat to human, livestock, and wildlife populations. In wildlife populations, disease has been responsible for dramatic declines, extirpations, and even extinctions in mammals, birds, echinoderms, amphibians, and other taxa [@hewson2014; @samuel2015; @scheele2019; @cunningham2021]. Although the dramatic and immediate nature of disease-induced declines in wildlife has led to intensive focus on documenting, understanding, and mitigating disease impacts [@langwig2015], populations previously decimated by disease are also beginning to recover [@newell2013; @voyles2018]. Given the nascent nature of wildlife recoveries in the presence of disease, we know far less about the mechanisms driving population recovery [@brannelly; @russell2020] and how these mechanisms can be leveraged to re-establish host populations across large spatial scales [@mendelson2019]. In particular, for species of conservation concern as a result of disease-induced declines, demonstrating the long-term viability and feasibility of potential recovery actions in the presence of disease and elucidating the mechanisms underlying long-term recovery success remain largely absent from conservation biology [@venesky2012; @mendelson2019].

@@

Given the dramatic impact of disease on global wildlife populations, reintroductions will often be necessary to reestablish extirpated populations at a landscape scale. However, there are several conditions that will often determine the feasibility of this much-needed recovery approach. First, populations that survived following disease arrival have evolved increased resistance. Second, increased resistance contributes to higher individual fitness in the presence of disease and allows positive population growth. Third, the fitness benefits gained through selection for increased resistance manifest beyond the natal environment, allowing for the establishment of populations reintroduced into novel environments. And fourth, reintroduced populations show long-term growth and persistence in a landscape of disease. To date, few of these conditions are documented for any single species impacted by disease, and as a result the reintroduction of resistant hosts as a means to recover species of conservation concern remains largely an untested idea [@scheele2021]. Furthermore, a lack of relevant information about disease or a lack of suitable options to overcome the threat of disease has resulted in the failure of numerous reintroduction efforts [@scheele2021; @fellers2007]. With the increase in emerging infectious diseases in recent decades [@daszak2000], these failures highlight the critical need to develop effective tools to recover disease-impacted biodiversity worldwide. <!--# Need to better tie these two sentences for the conditins listed above. -->

@@

During the period `r min(d1$year)`--`r max(d1$year)`, we conducted `r n_distinct(d1$site_id, d1$date)` translocations of adult frogs to `r n_distinct(d1$site_id)` recipient sites in Yosemite (`r transno_min`--`r transno_max` translocations per site; @fig-yosemap). Each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs collected from one of `r n_distinct(d1$donor)` donor populations (@fig-yosemap). Estimates of frog survival during the one year following translocation indicate that survival was highly variable between sites, but relatively constant within sites (for the subset of sites that received multiple translocations; @fig-translocation-survival). As such, for sites that received more than one translocation, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in all subsequent translocations to that site (median of posterior distribution = 4.14, uncertainty interval = 3.06\--5.28; see Supporting Information for analysis details). The relatively constant within-site frog survival across all translocations suggests an important effect of site characteristics on frog survival. <!--# How to extract these values from table produced by line 934 in translocation_survival_analysis.Rmd (or from associated code output)? -->

@@

@fig-exomemap

@@

![Map showing the locations of translocated and donor populations in Yosemite National Park (park boundary indicated by black polygon). To obscure the exact locations of populations, a small amount of random noise was added to all point coordinates. Elevations are indicated as in @fig-exomemap. See inset map in @fig-exomemap to locate Yosemite National Park within California. <!--# Move to Supporting Information. Add YOSE boundary to inset map of fig-exomemap? Also, is basemap necessary/helpful? Finally, better to add color coding showing which donor was used at which sites?  Mark W: I like the basemap.  -->](figures/map_translocation_points.png){#fig-yosemap width="450"}

![Median survival of each cohort of translocated frogs to the 12 recipient sites included in this study. Sites are arranged along the x-axis using the average of the median 1-year survival per translocation at each site. Frog survival was estimated using capture-mark-recapture methods. Dot colors indicate the donor population from which frogs in each translocated cohort were collected.<!--# Mark W: Do we want to include error bars on these median estimates?  -->](figures/translocation_survival_bysiteid.png){#fig-translocation-survival width="450"}

![For the best model identifying predictors of 1-year frog survival following translocation, estimated posterior density curves and shaded 95% uncertainty intervals for all predictor variables. Variables are considered important predictors if the associated uncertainty interval does not overlap zero (dashed red line). <!--# Move to Supporting Information. Change @fig-survival-postdens from mcmc_areas to mcmc_intervals? -->](figures/mcmc_areas_m1d.png){#fig-survival-postdens width="450"}

![Conditional effects of the important predictors of 1-year frog survival: (A) elevation of recipient site, (B) winter severity in the year following translocation, (C) sex, and (D) donor population. In (A) and (B), blue lines are medians and gray ribbons are 95% uncertainty intervals. In (C) and (D), box plots show medians, first and third quartiles, largest and smallest values within 1.5x interquartile range, and values outside the 1.5x interquartile range.](figures/cond_effects_plot.png){#fig-cond-effects}

@@

![Map showing the locations of Bd-naive or recovering frog populations from which DNA samples were collected for the exome capture analyses. Black polygons indicate the boundaries of Yosemite, Kings Canyon, and Sequoia National Parks (from top to bottom). To obscure the exact locations of populations, random noise was added to all point coordinates. Inset map shows the location of the study area within California. In both maps, elevation is indicated by the colored hillshade layer (dark green = lowest elevation, white = highest elevation). <!--# Remove this map. Map in Allie's fig-selectionresults is likely sufficient. -->](figures/map_exome_points.png){#fig-exomemap width="450"}

@@

![Allele frequencies of the outlier SNPs for each of the sampled naive and recovering *R. sierrae* populations. The results for RIN3 are consistent with strong selection on this gene following the arrival of Bd, and suggests are role for RIN3 in frog resistance against Bd infection and associated population recovery.](figures/outlier_map_draft_figure.png){#fig-allelefrequencies}

@@

DNA samples were collected via buccal swabbing [@broquet2007] of xx-xx frogs from seven Bd-naive populations and seven recovering populations. These populations represented a wide latitudinal range in the southern Sierra Nevada, from northern Yosemite National Park to south of Sequoia National Park (@fig-exomemap). To minimize potential biases associated with different latitudinal ranges of the two population types, we selected populations to represent similar latitudinal ranges.

@@

>>>>

Emerging infectious diseases pose a severe threat to human, livestock, and wildlife populations. In wildlife populations, disease has been responsible for dramatic declines, extirpations, and even extinctions in mammals, birds, echinoderms, amphibians, and other taxa [@hewson2014; @samuel2015; @scheele2019; @cunningham2021]. Although the dramatic and immediate nature of disease-induced declines in wildlife has led to intensive focus on documenting, understanding, and mitigating disease impacts [@langwig2015], populations previously decimated by disease are also beginning to recover [@newell2013; @voyles2018]. Given the nascent nature of wildlife recoveries in the presence of disease, we know far less about the mechanisms driving population recovery [@brannelly2021; @russell2020] and how these mechanisms can be leveraged to re-establish host populations across large spatial scales [@mendelson2019]. In particular, for species of conservation concern as a result of disease-induced declines, demonstrating the long-term viability and feasibility of potential recovery actions in the presence of disease and elucidating the mechanisms underlying long-term recovery success remain largely absent from conservation biology [@venesky2012; @mendelson2019].

@@

There are four conditions that will often determine the feasibility of this much-needed recovery approach. First, populations that survived following disease arrival have evolved increased resistance. Second, increased resistance contributes to higher individual fitness in the presence of disease and allows positive population growth. Third, the fitness benefits gained through selection for increased resistance manifest beyond the natal environment, allowing for the establishment of populations reintroduced into novel environments. And fourth, reintroduced populations show long-term growth and persistence in a landscape of disease. To date, few of these conditions are documented for any single species impacted by disease, and as a result the reintroduction of resistant hosts as a means to recover species of conservation concern remains largely an untested idea [@scheele2021]. Furthermore, a lack of relevant information about disease or a lack of suitable options to overcome the threat of disease has resulted in the failure of numerous reintroduction efforts [@scheele2021; @fellers2007]. With the increase in emerging infectious diseases in recent decades [@daszak2000], these failures highlight the critical need to develop effective tools to recover disease-impacted biodiversity worldwide. <!--# Need to better tie these two sentences to the conditions listed above. -->

@@

During the period `r min(d1$year)`--`r max(d1$year)`, we conducted `r n_distinct(d1$site_id, d1$date)` translocations of adult frogs to `r n_distinct(d1$site_id)` recipient sites in Yosemite (`r transno_min`--`r transno_max` translocations per site; @fig-yosemap) <!--# How to restart numbering for figures in SI, add "S" prefix? -->. Each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs collected from one of `r n_distinct(d1$donor)` donor populations (@fig-yosemap). Estimates of frog survival during the one year following translocation indicate that survival was highly variable between sites, but relatively constant within sites (for the subset of sites that received multiple translocations; @fig-translocation-survival). As such, for sites that received more than one translocation, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in all subsequent translocations to that site (median of posterior distribution = 4.14, uncertainty interval = 3.06\--5.28; see Supporting Information for analysis details). The relatively constant within-site frog survival across all translocations suggests an important effect of site characteristics on frog survival. <!--# How to extract these values from table produced by line 934 in translocation_survival_analysis.Rmd (or from associated code output)? -->

@@

@@

![Median survival of each cohort of translocated frogs to the 12 recipient sites included in this study. Sites are arranged along the x-axis using the average of the median 1-year survival per translocation at each site. Frog survival was estimated using capture-mark-recapture methods. Dot colors indicate the donor population from which frogs in each translocated cohort were collected. <!--# Include error bars on these median estimates. -->](figures/translocation_survival_bysiteid.png){#fig-translocation-survival width="450"}

![Conditional effects of the important predictors of 1-year frog survival: (A) elevation of recipient site, (B) winter severity in the year following translocation, (C) sex, and (D) donor population. In (A) and (B), blue lines are medians and gray ribbons are 95% uncertainty intervals. In (C) and (D), box plots show medians, first and third quartiles, largest and smallest values within 1.5x interquartile range, and values outside the 1.5x interquartile range. <!--# Switch C and D panels. -->](figures/cond_effects_plot.png){#fig-cond-effects}

@@

@@

![Allele frequencies of the outlier SNPs for each of the sampled naive and recovering *R. sierrae* populations. The results for RIN3 are consistent with strong selection on this gene following the arrival of Bd, and suggests a role for RIN3 in frog resistance against Bd infection and associated population recovery.](figures/outlier_map_draft_figure.png){#fig-allelefrequencies}

@@

DNA samples were collected via buccal swabbing [@broquet2007] of xx-xx frogs from seven Bd-naive populations and seven recovering populations. These populations represented a wide latitudinal range in the southern Sierra Nevada, from northern Yosemite National Park to south of Sequoia National Park (@fig-selectionresults). To minimize potential biases associated with different latitudinal ranges of the two population types, we selected populations to represent similar latitudinal ranges.

@@

![Map showing the locations of translocated and donor populations in Yosemite National Park (park boundary indicated by black polygon). To obscure the exact locations of populations, random noise was added to all point coordinates. Inset map shows the location of the study area within California. In both maps, elevation is indicated by the colored hillshade layer (dark green = lowest elevation, white = highest elevation). <!--# Add color coding showing which donor was used at which sites. Add inset map locating YOSE within California. Zoom in on eastern YOSE only instead of entire park. -->](figures/map_translocation_points.png){#fig-yosemap width="450"}

![For the best model identifying predictors of 1-year frog survival following translocation, estimated posterior density curves and shaded 95% uncertainty intervals for all predictor variables. Variables are considered important predictors if the associated uncertainty interval does not overlap zero (dashed red line).](figures/mcmc_areas_m1d.png){#fig-survival-postdens width="450"}
